---
title: "Lesson 2: Single Page R Markdown 📃 Document"
author: "Jarrod"
date: 2020-04-08T21:13:14-05:00
categories: ["R"]
#tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
library(DT)
library(Rcpp)
library(ggplot2)
library(dplyr)
library(blogdown)
library(htmltools)
```


{{% opennote %}}
1) [The Open Source Community](#the-open-source-community) For context.
2) [What you need help with](#what-is-your-goal).
Then I get into three general platorms for finding help.
3) [The Web](#the-web).
4) [Hosting Services like GitHub](#hosting-services).
5) [StackOverflow](#stackoverflow).
{{% /opennote %}}

R Markdown is a versitle framework for creating a range of [high-quality documents](https://rmarkdown.rstudio.com/gallery.html) including HTML, PDF, presentations, books, websites, and more. Markdown allows documents to be simple and portable. R gives you the benefit of creating dynamic documents that contain your results (e.g., tables, graphics, and inline values). Results can be computed and rendered dynamically from R code and as such are more likely to be reproducible.

> In this lesson you will (1) learn about the key parts of an R Markdown document; (2) create a simple, single page R Markdown document; (3) use basic methods to tweak the behavior and look of the document.

At the end of the lesson, I provide a minimal example template for you to reference while you make your own page.

## Basic Structure

A basic, single page, R Markdown document is created with just three element: the YAML header, Markdown-formatted text, and R code. Technically, you do not need the R code; many of the pages on this site have no R code at all. But for our purposes we will incorporate R code as a *neccessary ingredient* so that you can experience the potential of R Markdown. Even if you don't know much R, I want you to incorporate some R code in your document. I am happy to provide you with the tools you need. Many of your classmates are expert R coders so please ask for help on the `#course-curation` Slack channel. Further customization can be achived with additional CSS (Cascading Style Sheets) and/or HTML (Hypertext Markup Language) code, both of which will be the subject of a future lesson. For this lesson we focus on the three core elements.

### The YAML header

At the top of the document you must included the YAML metadata. YAML (YAML Ain't Markup Language)

For a single page document like the one you will create in this lesson, the YAML metadata need to be enclosed in delimiters, specifically three dashes (---). The delimeters must be on their own lines.

```
---
YAML metadata
---
```
Between the delimeters you write the metadata.



### Markdown text


### R code




```{r, echo=FALSE}
blogdown::shortcode('edit', 'https://github.com/stri-con/data-curation/blob/master/content/post/lesson-2.Rmd')
```

```{r, echo=FALSE}
blogdown::shortcode('anchor')
```
